---
title: "pysdg in R"
output: html_document
---

# Tutorial: Hellinger Distance

**This tutorial demonstrates how to calculate the Hellinger distance between real and synthetic data.** It assumes that pysdg is already installed in a conda environment. The package reticulate is used to interface with Python. For respective installation instructions, please consult the pysdg documentation. 

**Hellinger distance** is a fidelity (or broad utility) metric that compares the real data distribution to the synthetic data distribution. It can be measured per variable (univariate) or across the entire record (multivariate). As it is a distance measurement, lower values translate to higher fidelity.

## Step 1: Activate the Conda Environment
For the purpose of this tutorial, we assume that your conda environment was named `pysd_dev`, so the first step is to activate it. Note that Step 1 is *not* required if you are working with a pre-defined Docker image that has the required environment including its dependencies pre-installed.

```{r}
library(reticulate) 
use_condaenv("pysdg_dev", required = TRUE) 
```

## Step 2: Import Core and Evaluation Functions
Relevant functions are imported from pysdg: 
* synth.load for standardized loading of datasets via pysdg, 
* synth.generate to generate a synthetic dataset and 
* synth.metrics to apply the fidelity metric Hellinger distance for evaluation

```{r}
synth.generate <- reticulate::import("pysdg.synth.generate") 
synth.load <- reticulate::import("pysdg.synth.load")
synth.metrics <- reticulate::import("pysdg.synth.metrics")
```

## Step 3: Load the Training Data
Loading data via the pysdg loading function requires setting a "dummy" generator. This can be any arbitrary generator, as the loading process is independent of the generator.

```{r}
current_file_path <- rstudioapi::getActiveDocumentContext()$path
raw_data_path <- paste0(dirname(current_file_path), "/raw_data.csv")
raw_info_path <- paste0(dirname(current_file_path), "/raw_info.json")
gen <- synth.generate$Generator("synthcity_bayesian_network")
train_data <- gen$load(raw_data_path, raw_info_path)
```

## Step 4: Generate a Synthetic Dataset
The detailed steps to generate synthetic datasets are described in tutorial 2.0 - basics. In the current tutorial, we generate a synthetic dataset using the generator as specified in Step 3. Note that you can also directly load a synthetic dataset from your directory.

```{r}
gen$train()
gen$gen(num_rows = nrow(train_data), num_synths = 1)
synths <- gen$unload()
synth_data <- synths[[1]]
```

## Step 5: Calculate the Univariate Hellinger Distance 
For the calculation of Hellinger distance make sure that the real and synthetic dataset are consistent. Loading them via pysdg is a way that ensures such consistency. We calculate the univariate Hellinger distance in this example. It outputs three values:

* hlngr_median: the median Hellinger distance across all univariate calculations
* hlngr_iqr: the corresponding IQR
* hlngr_stddev: the corresponding SD 
* hlngr_breakdown: the Hellinger distance per variable

```{r}
univar_hellinger <- synth.metrics$calc_univar_hellinger_distance(train_data, synth_data)
print(paste0("The median univariate Hellinger distance is ", univar_hellinger$hlngr_median))
```
